---
title: "Data Visualization"
author: "Alfredo Hernández"
subtitle: "What was wrong with the Challenger?"
output:
  html_notebook:
    highlight: pygments
    theme: cosmo
    toc: yes
  pdf_document:
    fig_height: 3.5
    fig_width: 9
    toc: yes
---

```{r message=FALSE}
library(tidyverse)
library(ggplot2)
```

# Logistic regression

$y_{i} = Bernoulli(\pi_{i}) = \begin{cases} 1 & \pi_{i} \\ 0 & 1 - \pi_{i} \end{cases}$, where $\pi_{i}$ is the probability of failiure.

$\log \dfrac{\pi_{i}}{1 - \pi_{i}} = \beta_{0} + \beta_{1} T_{i}$ (logit transformation)

We test $H_{0}: \beta_{0} = 0$ against $H_{1}: \beta_{0} \neq 0$ and $H_{0}: \beta_{1} = 0$ against $H_{1}: \beta_{1} \neq 0$.

```{r}
temp <- c(66,70,69,68,67,72,73,70,57,63,70,78,67,53,67,75,70,81,76,79,75,76,58)
y <- c(0,1,0,0,0,0,0,0,1,1,1,0,0,1,0,0,0,0,0,0,1,0,1)

logit <- glm(y~temp, family = binomial)

a <- summary(logit)
a
```

The probability of failiure at $T = 31 F$ ($-0.6 C$) would have been:
```{r}
exp(15.0429 -0.2322 * 31)/ (1 + exp(15.0429 -0.2322 * 31))
```



# Approximate test (maximum likelihood approach)

```{r}
zinttrue<-abs(a$coefficients[1,3])
ztemptrue<-abs(a$coefficients[2,3])
nr=10000 #number of rearrangements to be examined
zint=numeric(nr);ztemp=numeric(nr)

for (i in 1:nr){
	newy<- sample(y, 23)
	logit<-glm(newy~temp,family=binomial)
	a<-summary(logit)
	zint[i]<-abs(a$coefficients[1,3])
	ztemp[i]<-abs(a$coefficients[2,3])
}
```

```{r}
length(zint[zint >= zinttrue])/nr
length(ztemp[ztemp >= ztemptrue])/nr
```

We conclude that the temperature $T$ ($p$-value = `r length(ztemp[ztemp >= ztemptrue])/nr`) is an even more important variable.
