---
title: "Data Visualization"
author: "Alfredo Hernández"
subtitle: "Parametric and non-parametric bootstrap"
date: "25 October 2017"
output:
  html_notebook:
    highlight: pygments
    theme: cosmo
    toc: yes
  pdf_document:
    fig_height: 3.5
    fig_width: 9
    toc: yes
---

# The parametric bootstrap for censored data

## Type II censored data (WIP)

It occurs when a life test is terminated exactly when a pre-specified number of failures have occurred. The remaining units have not yet failed. If $n$ units were on test, and the pre-specified number of failures is $r \leq n$, then the test ends at $t_{r} =$ the time of the $r$-th failure. For instance, this sample was
censored after 7 failures and the remaining units were 3.

The maximum likelihood estimator of the failure time mean is:
$$ \hat{\theta} = \dfrac{\sum_{i=1}^{r} t_{(i) + (n - r) t_{(r)}}}{r} $$

```{r}
time<-c(125, 189, 210, 356, 468, 550, 610)
ftm<-(sum(time)+3*max(time))/7
ftm
```

### Parametric bootstrap replicates
```{r}
nsim <- 1000
ftmr <- numeric(nsim)

for(i in 1:nsim){
	x <- rexp(10,1/ftm)
	y <- sort(x)
	y <- y[1:7]
	ftmr[i]<-(sum(y)+3*max(y))/7
}
```
```{r}
#Bootstrap estimate:
mean(ftmr)
#Expected Bias:
mean(ftmr) - ftm
#Estimated standard error:
sd(ftmr)
# 95% confidence intervals
quantile(ftmr, probs = c(0.025, 0.975))
```
This leads to the interval $[254 , 1138] : 696 \pm 442$.


## Type I censored data

It occurs when a life test is terminated exactly for a fixed time $c$. The remaining units have not yet failed. $n$ units were on test, and $r$ units have failed at a time less or equal to $c$ (where r is less than or equal to $n$). For instance, now this sample corresponds to a life test experiment that was ended at `c = 612`, and the remaining units were 3.

The maximum likelihood estimator of the failure time mean is:
$$ \hat{\theta} = \dfrac{\sum_{i=1}^{r} t_{(i) + (n - r) c}}{r} = \frac{T}{r} $$

```{r}
time <- c(125, 189, 210, 356, 468, 550, 610)
c <- 612
ftm <- (sum(time) + 3*c)/7
ftm
```

### Parametric bootstrap replicates
```{r}
nsim<-1000
ftmr<-numeric(nsim);
for(i in 1:nsim) {
	x<-rexp(10,1/ftm)
	y<-x[x<=c]
	r<-length(y)
	ftmr[i]<-(sum(y)+(10-r)*c)/r
}
```

```{r}
#Bootstrap estimate:
mean(ftmr)
#Expected Bias:
mean(ftmr) - ftm
#Estimated standard error:
sd(ftmr)
# 95% confidence intervals
quantile(ftmr, probs = c(0.025, 0.975))
```

# The non-parametric bootstrap

TODO: stuff (before page 37)

## CLESS: Common language effect size statistic

CLESS is an estimation of the probability that a randomly selected value from the population with the greatest expectation will be greater than a randomly sampled value from the other population.

$$ CLESS \approx P(X_{d} - X_{h} > 0) = P\left(Z < \frac{\mu_{d} - \mu_{h}}{\sigma \sqrt{2}}  \right) = \Phi(d/\sqrt{2}) $$

where $d$ is the effect size.

### Can normality be assumed?

The quotient 2D:4D is in some sense arbitrary. Why not consider $4D:2D$ instead?

It is natural to expect that the same statistical model used for $2D:4D$ could also describe the $4D:2D$ ratios, that is, the data set of the reciprocals.

However, if the r.v. $X$ is normally distributed $1/X$ is not normally distributed !!

We are going to analyze our data set using several methods:

```{r}
iw<-c(0.998,0.950,1.023,0.980,0.907,0.958,1.016,0.957,0.998,0.979,0.982,1.000,0.998,1.017,0.976,0.967,1.023,1.033,
0.979,0.983,0.990,0.998)
im<-c(0.931,1.019,0.936,0.974,0.949,0.939,0.984,0.956,0.947,0.969,0.979,0.922,0.973,1.014,0.936,0.966,1.020,0.975,1.015)
t.test(iw,im) # Assuming normality
```

Three kind of bootstrap CI for the difference of means:
```{r}
nw<-length(iw); 
nm<-length(im);
nb<-1000

difb<-numeric(nb)
for(i in 1:nb){
	iwb<-sample(iw,nw,replace=T)
	imb<-sample(im,nm,replace=T)
	difb[i]<-mean(iwb)-mean(imb)
}

#Quantile method:
quantile(difb,c(0.025,0.975))

# Simple method:
2*(mean(iw)-mean(im))-quantile(difb,0.975)
2*(mean(iw)-mean(im))-quantile(difb,0.025)
```

Bootstrap-t CI method:

$$ V(\bar{X}_{w} - \bar{X}_{m} ) = V(\bar{X}_{w}) + V(\bar{X}_{m}) = \frac{\sigma^{2}_{w}}{n_{w}} + \frac{\sigma^{2}_{m}}{n_{m}}$$
```{r}
tb <- numeric(nb)
theta <- mean(iw) - mean(im)
sdtheta <- sqrt(var(iw)/nw+var(im)/nm)

for(i in 1:nb){
	iwb<-sample(iw,nw,replace=T)
	imb<-sample(im,nm,replace=T)
	thetab<-mean(iwb)-mean(imb)
	sdthetab<- sqrt(var(iwb)/nw+var(imb)/nm)
	tb[i]<-(thetab-theta)/sdthetab
}

# Quantile method
quantile(tb,c(0.025,0.975))

# Simple method
theta+sdtheta*quantile(tb,0.025)
theta+sdtheta*quantile(tb,0.975)
```

## Non-linear regression model
```{r}
x <- c(77.6,114.9,141.1,190.8,239.9,289,332.8,378.4,
			 434.8,477.3,536.8,593.1,689.1,760)
y <-c (10.07,14.73,17.94,23.93,29.61,35.18,40.02,
			 44.82,50.76,55.05,61.01,66.4,75.47,81.78)

ajust<- nls(y ~ b1 * (1-exp(-b2*x)), start = list(b1= 2, b2=0.001))
summary(ajust)
```
```{r}
xx<-seq(1,10000)
pred<-summary(ajust)$coefficients[1] * (1-exp(-summary(ajust)$coefficients[2] *xx))
plot(xx,pred)
```

 STUFF


## The bootstrap method in AR(p) time series

WIP: THEORY

Consider the following time series with 30 observations:

```{r}
x<-c(12.1,12.9,17.5,12.3,14.3,13.1,12.5,10,10.9,11.7,13.2,
11.1,12.1,11.4,11.2,11.7,13.2,16.5,15.9,16.6,9.5,9,11.6,10.3,
12.4,10.4,12.9,19.1,12.1,8.5)
ts.plot(x)
```


We want to fit it using an $AR(2)$ model.

```{r}
fit <- ar(x, aic = FALSE, order.max = 2, method = "yule-walker")
fit
mean(x)
```

Fibbonacci's sequence is an example of a classical deterministic (i.e., there's no $Z_{t}$ noise) auto-regressive model:
$$ Y_{t} = Y_{t-1} + Y_{t-2} $$

#### Bootstraping the residuals
 1. Fit the original time series and keep the residuals and the
estimates of the parameters.
```{r}
res<-fit$resid
res
```
Note that the two first values are missing.


 2. Construct the bootstrap series using the estimated process
and a sample of the residuals instead of $Z_{t}$:
$$ YP_{t} = 11.554 + 0.287 YP_{t-1} - 0.209 YP_{t-2} + resp_{t} $$

```{r}
nb<-1000; mu<-numeric(nb); phi1<-numeric(nb)
phi2<-numeric(nb)
xb<-numeric(30)
for(j in 1:nb){
xb[1]<-x[1];xb[2]<-x[2]
for (i in 3:30){
r<-sample(res[3:30],1,replace=T)
xb[i]<-fit$ar[1]*xb[i-1]+fit$ar[2]*xb[i-2]+mean(x)*(1-fit$ar[1]-
fit$ar[2])+r}
fitb<-ar(xb, aic = FALSE, order.max = 2,method=c("yule-walker"))
mu[j]<-mean(xb);phi1[j]<-fitb$ar[1]; phi2[j]<-fitb$ar[2] }
```

