---
title: "Data Visualization"
author: "Alfredo Hernández"
subtitle: Exercise 1
output:
  html_notebook:
    highlight: pygments
    theme: cosmo
    toc: yes
  pdf_document:
    toc: yes
---

# Exercise 1.1

Remember one of the examples presented in class: The first row is chest circumference (in inches) of five subjects. Let us call this $X$. The second row is the respective total volumes of air that can be breathed in and out in one minute (in liters) for the same five subjects. Let us call
this $Y$.

```{r}
X <- c(39, 29, 60, 40, 32)
Y <- c(11, 5, 20, 8, 6)
```
- Download the package `combinat` in order to use the function `permn()`. Perform an exact test for $H_{0} : \rho = 0$ against $H_{1} : \rho > 0$. Use Pearson and Spearman correlation coefficient as well.
```{r message=FALSE}
library(combinat)
library(stats)
library(ggplot2)
```

```{r}
n.perm.Y <- fact(length(Y))
rho.true.p <- cor(X,Y, method = "pearson")
rho.true.s <- cor(X,Y, method = "spearman")
```


```{r}
perm.Y <- permn(Y)
```
## Pearson Correlation Coefficient
```{r}
# Pearson correlation coefficient
rho.p <- numeric(n.perm.Y)
count <- 0
for (i in 1:n.perm.Y) {
	rho.p[i] <- cor(X, perm.Y[[i]])
	if (rho.p[i] > rho.true.p) {
		count <- count + 1
	}
}
p.val.p <- count/n.perm.Y
p.val.p
```

```{r}
# Histogram of rho
ggplot() +
	geom_histogram(aes(rho.p), colour="black", fill="white") +
	geom_vline(aes(xintercept = quantile(rho.p, prob = 1 - p.val.p), colour = "1 - p-value"), 
						 show.legend = T) +
	geom_vline(aes(xintercept = quantile(rho.p, prob = 0.95), colour = "1 - alpha"), 
						 linetype = "longdash", show.legend = T) +
	labs(title = paste0("Cor. coeff. histogram; p-value = ", format(round(p.val.p, 4), nsmall = 4), " (null hypothesis ", ifelse(p.val.p < 1-0.95 ,"rejected", "failed to be rejected"), ")"),
			 colour = "quantile") +
	scale_color_manual(values = c("red","blue"))
```

## Spearman Correlation Coefficient
```{r}
# Spearman correlation coefficient
rho.s <- numeric(n.perm.Y)
count <- 0
for (i in 1:n.perm.Y) {
	rho.s[i] <- cor(X, perm.Y[[i]], method = "spearman")
	if (rho.s[i] > rho.true.s) {
		count <- count + 1
	}
}
p.val.s <- count/n.perm.Y
p.val.s
```

```{r}
# Histogram of rho
ggplot() +
	geom_histogram(aes(rho.s), colour="black", fill="white") +
	geom_vline(aes(xintercept = quantile(rho.s, prob = 1 - p.val.s), colour = "1 - p-value"), 
						 show.legend = T) +
	geom_vline(aes(xintercept = quantile(rho.s, prob = 0.95), colour = "1 - alpha"), 
						 linetype = "longdash", show.legend = T) +
	labs(title = paste0("Cor. coeff. histogram; p-value = ", format(round(p.val.p, 4), nsmall = 4), " (null hypothesis ", ifelse(p.val.s < 1-0.95 ,"rejected", "failed to be rejected"), ")"),
			 colour = "quantile") +
	scale_color_manual(values = c("red","blue"))
```

# Exercise 1.2

These are the increments of weight recorded in an experiment where a new additive has been added to a standard compound feed:

```{r}
weight.std <- c(2.5, 3.4, 2.9, 4.1, 5.3, 3.4, 1.9, 3.3, 1.8)
weight.add <- c(3.5, 6.3, 4.2, 4.3, 3.8, 5.7, 4.4)
```

## Classical T-test
- Is the additive efficient? Perform a classical T-test and use the
corresponding T-statistic to perform a permutation test.

Here we are testing $H_{0} : \operatorname{E}(add) = \operatorname{E}(std)$ against $H_{1} : \operatorname{E}(add) > \operatorname{E}(std)$.

```{r}
n.re <- 10000
n.std <- length(weight.std)
n.add <- length(weight.add)
n.total <- n.std + n.add
```

```{r}
st.true <- mean(weight.add) - mean(weight.std)
st <- numeric(n.re)

count <- 0
weight.all <-  c(weight.std, weight.add)

for (i in 1:n.re) {
	d <- sample(weight.all, n.total)
	test.add <- d[1:n.add]
	a <- n.add + 1
	test.std <- d[a:n.total]
	st[i] <- mean(test.add) - mean(test.std)
	if (st[i] > st.true) {
		count <- count + 1
	}
}
p.val <- count/n.re
p.val
```

```{r}
# Histogram of mean statistics
ggplot() +
	geom_histogram(aes(st), colour="black", fill="white") +
	geom_vline(aes(xintercept = quantile(st, prob = 1 - p.val), colour = "1 - p-value"), 
						 show.legend = T) +
	geom_vline(aes(xintercept = quantile(st, prob = 0.95), colour = "1 - alpha"), 
						 linetype = "longdash", show.legend = T) +
	labs(title = paste0("mean statistics histogram; p-value = ", format(round(p.val, 4), nsmall = 4), " (null hypothesis ", ifelse(p.val.s < 1-0.95 ,"rejected", "failed to be rejected"), ")"),
			 colour = "quantile") +
	scale_color_manual(values = c("red","blue", "green"))
```

## Other statistics
- Explore several statistics in order to perform the permutation
test. What is the final conclusion?

Here we are testing $H_{0} : \operatorname{median}(add) = \operatorname{median}(std)$ against $H_{1} : \operatorname{median}(add) > \operatorname{median}(std)$.

```{r}
st.true <- median(weight.add) - median(weight.std)
st <- numeric(n.re)

count <- 0
weight.all <-  c(weight.std, weight.add)

for (i in 1:n.re) {
	d <- sample(weight.all, n.total)
	test.add <- d[1:n.add]
	a <- n.add + 1
	test.std <- d[a:n.total]
	st[i] <- median(test.add) - median(test.std)
	if (st[i] > st.true) {
		count <- count + 1
	}
}
p.val <- count/n.re
p.val
```

```{r}
# Histogram of median statistics
ggplot() +
	geom_histogram(aes(st), colour="black", fill="white") +
	geom_vline(aes(xintercept = quantile(st, prob = 1 - p.val), colour = "1 - p-value"), 
						 show.legend = T) +
	geom_vline(aes(xintercept = quantile(st, prob = 0.95), colour = "1 - alpha"), 
						 linetype = "longdash", show.legend = T) +
	labs(title = paste0("median statistics histogram; p-value = ", format(round(p.val, 4), nsmall = 4), " (null hypothesis ", ifelse(p.val.s < 1-0.95 ,"rejected", "failed to be rejected"), ")"),
			 colour = "quantile") +
	scale_color_manual(values = c("red","blue", "green"))
```

### Conclusions
We conclude that, indeed, the null hypothesis $H_{0}: \verb|weight.add| = \verb|weight.std|$ is rejected; which in turn tells us they are different distributions.
