---
title: "Data Visualization"
author: "Alfredo Hernández"
subtitle: Exercise 1
output:
  pdf_document:
    fig_height: 3.5
    fig_width: 9
    toc: yes
  html_notebook:
    highlight: pygments
    theme: cosmo
    toc: yes
---

# Exercise 1.1

Remember one of the examples presented in class: The first row is chest circumference (in inches) of five subjects. Let us call this $X$. The second row is the respective total volumes of air that can be breathed in and out in one minute (in liters) for the same five subjects. Let us call
this $Y$.

```{r}
X <- c(39, 29, 60, 40, 32)
Y <- c(11, 5, 20, 8, 6)
```
- Download the package `combinat` in order to use the function `permn()`.
- Perform an exact test for $H_{0} : \rho = 0$ against $H_{1} : \rho > 0$. Use Pearson and Spearman correlation coefficient as well.
```{r message=FALSE}
# Necessary libraries
library(combinat)
library(stats)
library(ggplot2)
```

```{r}
n.perm.Y <- fact(length(Y))
rho.true.p <- cor(X,Y, method = "pearson")
rho.true.s <- cor(X,Y, method = "spearman")
```

```{r}
perm.Y <- permn(Y)
```
## Pearson Correlation Coefficient
```{r}
# Pearson correlation coefficient
rho.p <- numeric(n.perm.Y)
count <- 0
for (i in 1:n.perm.Y) {
  rho.p[i] <- cor(X, perm.Y[[i]], method = "pearson")
  if (rho.p[i] > rho.true.p) {
    count <- count + 1
  }
}
p.val.p <- count/n.perm.Y
p.val.p
```

```{r message=FALSE}
# Histogram of rho
ggplot() +
  geom_histogram(aes(rho.p), colour="black", fill="white") +
  geom_vline(aes(xintercept = quantile(rho.p, prob = 1 - p.val.p), 
                 colour = "1 - p-value"), show.legend = T) +
  geom_vline(aes(xintercept = quantile(rho.p, prob = 0.95), 
                 colour = "1 - alpha"), linetype = "longdash", show.legend = T) +
  labs(title = paste0("Cor. coeff. histogram; p-value = ",
                      format(round(p.val.p, 4), nsmall = 4), " (null hypothesis ", 
                      ifelse(p.val.p < 1-0.95 , "rejected", "failed to be rejected"), ")"),
       colour = "quantile") +
  scale_color_manual(values = c("red","blue"))
```

## Spearman Correlation Coefficient
```{r}
# Spearman correlation coefficient
rho.s <- numeric(n.perm.Y)
count <- 0
for (i in 1:n.perm.Y) {
  rho.s[i] <- cor(X, perm.Y[[i]], method = "spearman")
  if (rho.s[i] > rho.true.s) {
    count <- count + 1
  }
}
p.val.s <- count/n.perm.Y
p.val.s
```

```{r message=FALSE}
# Histogram of rho
ggplot() +
  geom_histogram(aes(rho.s), colour="black", fill="white") +
  geom_vline(aes(xintercept = quantile(rho.s, prob = 1 - p.val.s), 
             colour = "1 - p-value"), show.legend = T) +
  geom_vline(aes(xintercept = quantile(rho.s, prob = 0.95), 
             colour = "1 - alpha"), linetype = "longdash", show.legend = T) +
  labs(title = paste0("Cor. coeff. histogram; p-value = ",
                      format(round(p.val.p, 4), nsmall = 4), " (null hypothesis ", 
                      ifelse(p.val.s < 1-0.95 , "rejected", "failed to be rejected"), ")"),
       colour = "quantile") +
  scale_color_manual(values = c("red","blue"))
```

# Exercise 1.2

These are the increments of weight recorded in an experiment where a new additive has been added to a standard compound feed:

```{r}
weight.std <- c(2.5, 3.4, 2.9, 4.1, 5.3, 3.4, 1.9, 3.3, 1.8)
weight.add <- c(3.5, 6.3, 4.2, 4.3, 3.8, 5.7, 4.4)
```

```{r}
# Values used in some of the different tests
n.re <- 10000
n.std <- length(weight.std)
n.add <- length(weight.add)
n.total <- n.std + n.add
```

## Classical T-test
- Is the additive efficient? Perform a classical T-test and use the
corresponding T-statistic to perform a permutation test.

Here we are testing $H_{0} : \operatorname{E}(add) = \operatorname{E}(std)$ against $H_{1} : \operatorname{E}(add) > \operatorname{E}(std)$.

```{r}
# Mean statistics
mn.true <- mean(weight.add) - mean(weight.std)
mn <- numeric(n.re)

count <- 0
weight.all <- c(weight.std, weight.add)

for (i in 1:n.re) {
  d <- sample(weight.all, n.total)
  test.add <- d[1:n.add]
  a <- n.add + 1
  test.std <- d[a:n.total]
  mn[i] <- mean(test.add) - mean(test.std)
  if (mn[i] > mn.true) {
    count <- count + 1
  }
}
p.val.mn <- count/n.re
p.val.mn
```

```{r message=FALSE}
# Histogram of mean statistics
ggplot() +
  geom_histogram(aes(mn), colour="black", fill="white") +
  geom_vline(aes(xintercept = quantile(mn, prob = 1 - p.val.mn), 
             colour = "1 - p-value"), show.legend = T) +
  geom_vline(aes(xintercept = quantile(mn, prob = 0.95), 
             colour = "1 - alpha"), linetype = "longdash", show.legend = T) +
  labs(title = paste0("mean statistics histogram; p-value = ",
                      format(round(p.val.mn, 4), nsmall = 4), " (null hypothesis ", 
                      ifelse(p.val.mn < 1-0.95 , "rejected", "failed to be rejected"), ")"),
       colour = "quantile") +
  scale_color_manual(values = c("red","blue", "green"))
```

## Other statistics
- Explore several statistics in order to perform the permutation
test. What is the final conclusion?

### Median statistics
Here we are testing $H_{0} : \operatorname{median}(add) = \operatorname{median}(std)$ against $H_{1} : \operatorname{median}(add) > \operatorname{median}(std)$.

```{r}
# Median statistics
mdn.true <- median(weight.add) - median(weight.std)
mdn <- numeric(n.re)

count <- 0
weight.all <-  c(weight.std, weight.add)

for (i in 1:n.re) {
  d <- sample(weight.all, n.total)
  test.add <- d[1:n.add]
  a <- n.add + 1
  test.std <- d[a:n.total]
  mdn[i] <- median(test.add) - median(test.std)
  if (mdn[i] > mdn.true) {
    count <- count + 1
  }
}
p.val.mdn <- count/n.re
p.val.mdn
```

```{r message=FALSE}
# Histogram of median statistics
ggplot() +
  geom_histogram(aes(mdn), colour="black", fill="white") +
  geom_vline(aes(xintercept = quantile(mdn, prob = 1 - p.val.mdn), 
             colour = "1 - p-value"), show.legend = T) +
  geom_vline(aes(xintercept = quantile(mdn, prob = 0.95), 
             colour = "1 - alpha"), linetype = "longdash", show.legend = T) +
  labs(title = paste0("median statistics histogram; p-value = ",
                      format(round(p.val.mdn, 4), nsmall = 4), " (null hypothesis ", 
                      ifelse(p.val.mdn < 1-0.95 , "rejected", "failed to be rejected"), ")"),
       colour = "quantile") +
  scale_color_manual(values = c("red","blue", "green"))
```

### Proportion statistics
Here we are testing $H_{0} : \operatorname{E}(add)/\operatorname{E}(std) = 1$ against $H_{1} : \operatorname{obs}(add)/\operatorname{obs}(std) > 1$, where the permutations will give us an exact test (we understand $\operatorname{obs}(X)$ as an observation from the $X$ variable).

```{r}
# Proportion statistics
prop.true <- mean(weight.add)/mean(weight.std)
prop <- numeric(n.add * n.std)

count <- 0
k <- 1

for (i in 1:n.std) {
  for (j in 1:n.add) {
    prop[k] <- weight.add[j]/weight.std[i]
    k <- k + 1
  }
}

for (k in length(prop)) {
  if (prop[k] > prop.true) {
    count <- count + 1
  }
}

p.val.prop <- count/n.re
p.val.prop
```

```{r message=FALSE}
# Histogram of proportion statistics
ggplot() +
  geom_histogram(aes(prop), colour="black", fill="white") +
  geom_vline(aes(xintercept = quantile(prop, prob = 1 - p.val.prop), 
             colour = "1 - p-value"), show.legend = T) +
  geom_vline(aes(xintercept = quantile(prop, prob = 0.95), 
             colour = "1 - alpha"), linetype = "longdash", show.legend = T) +
  labs(title = paste0("proportion statistics histogram; p-value = ",
                      format(round(p.val.prop, 4), nsmall = 4), " (null hypothesis ", 
                      ifelse(p.val.prop < 1-0.95 , "rejected", "failed to be rejected"), ")"),
       colour = "quantile") +
  scale_color_manual(values = c("red","blue", "green"))
```

## Conclusions
We conclude that, indeed, the null hypothesis $H_{0}: \verb|weight.add| = \verb|weight.std|$ is rejected; which in turn tells us they are different distributions, and that indeed (given the design of our different tests) the new additive is efficient.
